---
title: "CHESS power"
author: "Paul Conn"
date: "5/12/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




We conducted a simple Monte Carlo simulation study to investigate the power of two aerial
surveys to detect a population decline in bearded and ringed seals in the Chukchi Sea.  We assumed two
consecutive surveys would have similar precision levels to that achieved in the present survey (CV=0.13 for bearded seals and CV=0.11 for ringed seals), and considered three different levels of population decline: 10\%, 25\%, and 50\%, we   
repeatedly simulated population estimates (assumed to have a lognormal distribution) and 
calculated the proportion of simulation replicates that
a 95\% two-sided z-test (conducted on the log scale) correctly suggested a population decline.

```{r sims}
# power to detect a decline
Decline = c(0.1,0.25,0.5)
CV=c(.13,.11)
n_reps = 1000
N1 = 100000  #this shouldn't really matter since we're using CV
Boot_out = array(0,dim=c(3,2,n_reps)) #array dims hold decline, species, sim replicate 


for(idec in 1:3){
  for(isp in 1:2){
  for(irep in 1:n_reps){
    N2 = (1-Decline[idec])*N1
    log_N1 = log(N1)
    log_N2 = log(N2)
    sigma = sqrt(log((CV[isp])^2+1))
    logN_est1 = rnorm(1,log_N1,sigma)
    logN_est2 = rnorm(1,log_N2,sigma)
    sigma_diff = sqrt(2)*sigma
    test_stat = (logN_est1-logN_est2)/sigma_diff
    #makes more sense to do z-test on log scale since N assumed lognormally distributed
    if(test_stat>1.96)Boot_out[idec,isp,irep]=1 #critical value for 2-sided z-test 
  }
  }
}
```

This short study suggested that for bearded seals, that 10\% declines
would be detected `r 100*sum(Boot_out[1,1,])/n_reps` percent of the time, 25\% declines would be detected `r 100*sum(Boot_out[2,1,])/n_reps` percent of the time, and 50\% declines would be detected `r 100*sum(Boot_out[3,1,])/n_reps` percent of the time.  For ringed seals, these percentages were `r 100*sum(Boot_out[1,2,])/n_reps`, `r 100*sum(Boot_out[2,2,])/n_reps`, and `r 100*sum(Boot_out[3,2,])/n_reps`, respectively.
